---
title: "Maps"
output: html_notebook
---

## Processing UFO data

Loading in libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(tidygeocoder)
```

Reading in US data

```{r message=FALSE, warning=FALSE}
ufo_data_us <- read_csv('../data/ufo_data_us_cleaned_v2.csv')
```

Some cleaning

```{r}
# Converting time posted
ufo_data_us$Posted <- as.Date(ufo_data_us$Posted, format = "%m/%d/%y")

# Clean time of sighting (only saves Dates)
ufo_data_us$`Date/Time` <- format(as.Date(ufo_data_us$`Date/Time`, format = "%m/%d/%y"), "%Y")
ufo_data_us$`Date/Time` <- as.numeric(ufo_data_us$`Date/Time`)
```

Aggregating UFO data for historical time series (wide format to match stacked D3 format)

```{r}
historical_data <- ufo_data_us %>%
  mutate(`Date/Time` = case_when(
          `Date/Time` > 2021 ~ `Date/Time` - 100,
          TRUE ~ `Date/Time`)
        ) %>% 
  group_by(`Date/Time`) %>% 
  summarize(
    Total = n()
  ) %>% 
  drop_na() 

# Get UFO data by state
states <- ufo_data_us %>% 
  mutate(`Date/Time` = case_when(
          `Date/Time` > 2021 ~ `Date/Time` - 100,
          TRUE ~ `Date/Time`)
        ) %>% 
  group_by(`Date/Time`, State) %>% 
  summarize(
    Sightings = n(),
    .groups = 'keep'
  ) %>% 
  drop_na() %>% 
  pivot_wider(names_from = State, values_from = Sightings, values_fill = list(Sightings = 0))

# Get UFO data by shape
shapes <- ufo_data_us %>% 
  mutate(`Date/Time` = case_when(
          `Date/Time` > 2021 ~ `Date/Time` - 100,
          TRUE ~ `Date/Time`)
        ) %>% 
  group_by(`Date/Time`, Shape) %>% 
  summarize(
    Sightings = n(),
    .groups = 'keep'
  ) %>% 
  drop_na() %>% 
  pivot_wider(names_from = Shape, values_from = Sightings, values_fill = list(Sightings = 0))

# Wide format joined
wide_joined <- inner_join(states, shapes, by = 'Date/Time')
wide_joined <- inner_join(wide_joined, historical_data, by = 'Date/Time')

# Writes csv
write_csv(wide_joined, '../webpage/data/us-historical-wide.csv')
```

Merging routes and UFO data

Reading in flight data

```{r}
airport <- read.csv('../data/airports.csv')
routes <- read.csv('../data/routes.csv')
```

Reverse geocoding to retrieve states by longitude and latitude

(Note reverse geocoding may take a while to query the results)

```{r}
airport_us <- airport %>% filter(Country == 'United States')

airport_us <- reverse_geocode(airport_us, lat = Latitude, long = Longitude, full_results = TRUE)
```

Selecting needed columns

```{r}
airport_us_1 <- airport_us %>% 
  select(Airport.ID, Name, City, state) %>% 
  rename(Source_Name = Name, Source_City = City,
         Source_State = state)

airport_us_2 <- airport_us %>% 
  select(Airport.ID, Name, City, state) %>% 
  rename(Dest_Name = Name, Dest_City = City,
         Dest_State = state)

routes <- routes %>% select(Source.airport.ID, Destination.airport.ID)
```

Converting state name to abbreviations

```{r}
names(state.abb) <- state.name 
airport_us_new <- airport_us
airport_us_new$State <- state.abb[airport_us_new$state]
```

Getting airport count by state and city

```{r}
airport_us_city <- airport_us_new %>% 
  select(Airport.ID, Name, City, State) %>% 
  group_by(City, State) %>% 
  summarize(
    `Airport Count` = n(),
    .groups = 'keep'
  ) %>% 
  arrange(desc(`Airport Count`))

airport_us_state <- airport_us_new %>% 
  select(Airport.ID, Name, City, State) %>% 
  group_by(State) %>% 
  summarize(
    `Airport Count` = n(),
    .groups = 'keep'
  ) %>% 
  arrange(desc(`Airport Count`))
```

Getting alien count by state and city

```{r}
ufo_us_city <- ufo_data_us %>% 
  group_by(City, State) %>% 
  summarize(
    Sightings = n(),
    .groups = 'keep'
  ) %>% 
  arrange(desc(Sightings))

ufo_us_state <- ufo_data_us %>% 
  group_by(State) %>% 
  summarize(
    Sightings = n()
  ) %>% 
  arrange(desc(Sightings))
```

Joining airport and sightings

```{r}
airport_and_ufo_city <- ufo_us_city %>% 
  left_join(airport_us_city, by = c("City", "State"))

airport_and_ufo_city$`Airport Count`[is.na(airport_and_ufo_city$`Airport Count`)] <- 0
  
airport_and_ufo_state <- ufo_us_state %>%
  left_join(airport_us_state, by = "State")

airport_and_ufo_state$`Airport Count`[is.na(airport_and_ufo_state$`Airport Count`)] <- 0
```

Checking 

```{r}
airport_and_ufo_state %>% 
  ggplot() +
  aes(Sightings) + 
  geom_histogram()
```

```{r}
airport_and_ufo_city %>% 
  ggplot() +
  aes(`Airport Count`) + 
  geom_histogram()
```


```{r}
airport_and_ufo_state %>% 
  ggplot() +
  aes(Sightings, `Airport Count`) +
  geom_point() +
  geom_smooth()
```

```{r}
airport_and_ufo_city %>% 
  ggplot() +
  aes(Sightings, `Airport Count`) +
  geom_point() +
  geom_smooth()
```

```{r}
airport_and_ufo_city %>% 
  ggplot() +
  aes(`Airport Count`, Sightings, group = as.factor(`Airport Count`)) + 
  geom_boxplot()
```

```{r}
write.csv(airport_and_ufo_city, '../webpage/data/airports-and-sightings-city.csv')
write.csv(airport_and_ufo_state, '../webpage/data/airports-and-sightings-state.csv')
```

